Molecular dynamics simulation of synchronization in driven particles

### 2020-11-10 manuscript/manuscript_ajp_2020.tex
updated text with unit definitions

\section{Introduction} %
%
%Outline
%-----------------------------------------------------------
%describe types of physical systems - particle types
Numerical simulations of confined, driven particles
can be used to model a variety of physical phenomena.
Particles which interact over long distances
include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [more detail and references]
Particles which interact over short distances include
bubble arrays/emulsions [more systems and references].

%------------------------------------------------------------------
%synchronization
Synchronization is a universal phenomena. 
Christiaan Huygens is credited with its discovery
with his 1665 observation of
in-phase oscillation of pendulum clocks \cite{Pikovsky2003}.  
Huygens explanation
that the clocks vibrations are coupled through the supporting
wall has since been rigorously demonstrated by 20th century
researchers \cite{Ramirez2016}.
Examples of synchronization occur
in many naturally oscillating systems,
from fireflies syncing with a pulsed LED \cite{},
to simultaneous flapping of birds wings \cite{},
to neuronal firing cardiac muscle \cite{} and brain tissue \cite{}.

%phase locking
Dynamical mode locking is a controlled 
synchronization phenomena in 
which an external driving force
causes the syncing of natural oscillation frequencies
in
weakly coupled oscillators.
Mode locking is often observed in quantum electronic
devices such as Josephson junctions \cite{}
and phase-locked loops \cite{}.
Instead of ohmic current-voltage (IV) relationships,
such devices exhibit stepped regions
such as Shapiro steps \cite{} [explain].

@@

%----------------------------------------------------------------------------
%paper outline
In the following paper we describe
our molecular dynamics model in Section~\ref{sec:MD}.
%including various  and confining environments.
We include code to simulate
and visualize the dynamics in this section
and supplementary material.
In Section~\ref{sec:results} we summarize
our results,
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one} and
multiple interacting particles
in Section~\ref{sec:sync},
including stationary propagation of high density kinks
in Sec.~\ref{sec:kink}.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of particle velocity versus position.
In Section~\ref{sec:quasiperiod},
we show how an aperiodic landscape modifies the particle dynamics.
Finally we explore the transition to chaotic dynamics in 
Section~\ref{sec:chaos}
and conclude 
in Section~\ref{sec:conclusion}. 
In each section
we suggest exercises for interested students,
and summarize our suggestions in Section~\ref{sec:problems}.

@@

\section{Molecular Dynamics Model}
\label{sec:MD}
We use a classical two-dimensional model for 
studying the dynamics of $N$ interacting particles. 
Particles are confined in a two-dimensional (2D) 
simulation of area $A = L \times L$ where $L=36.5$.
An individual particle $i$ has
position $\vec{r}_i = x_i \hat{x} + y_i \hat{y}$.
Particles are subject to
periodic boundary conditions
such that a particle leaving the edges of the system is mapped
back to a position within the simulation 
by the rules $x_i+L=x_i$ and $y_i+L=y_i$.

>>>>

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
that can be modeled with numerical studies of confined,
driven particles.
Christiaan Huygens 1665 observation of
in-phase oscillation of pendulum clocks \cite{Pikovsky2003}
[has been adapted into many useful technologies...]
%Huygens explanation
%that the clocks vibrations are coupled through the supporting
%wall has since been rigorously demonstrated by 20th century
%researchers \cite{Ramirez2016}.
Examples of synchronization occur
in many naturally oscillating systems,
from fireflies signaling in time with a with a pulsed LED \cite{},
to simultaneous flapping of birds wings \cite{},
to neuronal firing in cardiac muscle \cite{} and brain tissue \cite{}.

%phase locking
Dynamical mode locking is a controlled 
synchronization phenomena in 
which an external driving force
causes
%the natural oscillation frequencies
%in
weakly coupled oscillators to come into phase.
Mode locking is often observed in quantum electronic
devices such as Josephson junctions \cite{}
and phase-locked current loops \cite{}.
Instead of ohmic current-voltage (IV) relationships,
such devices exhibit stepped regions
such as Shapiro steps \cite{} [explain].

@@

%-----------------------------------------------------------
%describe types of physical systems - particle types
Numerical simulations of confined, driven particles
can be used to model a variety of physical phenomena.
Particles which interact over long distances
include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [more detail and references]
Particles which interact over short distances include
bubble arrays/emulsions [more systems and references].

%----------------------------------------------------------------------------
%paper outline
In the following paper we describe
our molecular dynamics model in Section~\ref{sec:MD}.
%including various  and confining environments.
We include code to simulate
and visualize the dynamics in this section
and supplementary material.
In Section~\ref{sec:results} we summarize
our results,
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one} and
multiple interacting particles
in Section~\ref{sec:sync},
including stationary propagation of high density kinks
in Sec.~\ref{sec:kink}.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of particle velocity versus position.
In Section~\ref{sec:quasiperiod},
we show how an aperiodic landscape modifies the particle dynamics.
Finally we explore the transition to chaotic dynamics in 
Section~\ref{sec:chaos}
and conclude 
in Section~\ref{sec:conclusion}. 
In each section
we suggest exercises for interested students,
as summarized in Section~\ref{sec:problems}.

@@

\section{Molecular Dynamics Model}
\label{sec:MD}
We use a classical two-dimensional model for 
studying the dynamics of $N$ interacting particles. 
Particles are confined in a two-dimensional (2D) 
simulation of area $A = L \times L$ where $L=36.5 a_0$
where $a_0$ is a dimensionless unit of length.
An individual particle $i$ has
position $\vec{r}_i = x_i \hat{x} + y_i \hat{y}$.
Particles are subject to
periodic boundary conditions
such that a particle leaving the edges of the system is mapped
back to a position within the simulation 
by the rules $x_i+L=x_i$ and $y_i+L=y_i$.
We summarize unit units of our simulated variables in Table ~\ref{tab:1}.

\begin{table}[h!]
\centering
\caption{Simulation units defined in the code.}
\begin{ruledtabular}
\begin{tabular}{c c p{5cm}}
Quantity & Unit \\
\hline
time &  $\eta / E_0$ \\
length &  $a_0$ \\
velocity &  \\
force & $E_0 / a_0$\\
electric potential & V(r_{ij} = q^2 E_0/r_{ij}), $E_0 = Z^*^2/()4\pi \epsilon \epsilon_0 a_0$ \\
charge/dimensionless interaction strength $q$ \\
effective colloidal charge & $Z^*$ \\
solvent dielectic constant & $\epsilon$ \\
\end{tabular}
\end{ruledtabular}
\label{tab:1}
\end{table}
